### 2023-06-15 python_excel_ite.tex
intro needs work

@@

Created PCDA course in response. Challenges with finding good textbooks. Developed much of my own curricular materials. Taught using combo of R Markdown documents and Jupyter notebooks. Literate programming. PCDA course focused mostly on data wrangling, EDA, statistical and machine learning based predictive models though I did include a Python based simulation of the Monte Hall 3-door problem. In teaching EDA using tools like ggplot2, I would challenge students to think about how they would go about trying to do some of the plots we were making in Excel. There is also some limited coverage of using Python to open Excel files and do some sheet manipulation. However, there was no real attempt to do the kind of modeling that was the focus of textbooks like PMS or DADM.

@@

@@

There are two overarching learning objectives for this first set of Jupyter notebooks. Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet - model building and sensitivity analysis. At the same time, students encounter, as they are needed, more advanced Python programming techniques such as object oriented programming and manipulation of tuples, lists and dictionaries. In this way, the computational problem to be solved drives the introduction of more advanced Python concepts and techniques instead of such things being presented in a vacuum. I make it a point to try to convey the notion that there are often multiple possible approaches by embarking on one path and there can be value to starting over and doing things in a different way. Each notebook is accompanied by videos that walk the students through the material and act as a guided tour with challenges to overcome along the way. The screenshot in \ref{fig:activities} shows the portion of the course web page that provides the high level roadmap through the notebooks and associated screencasts.

@@

The stage is set with an overview of a typical spreadsheet based model.

@@

The example used throughout the rest of the notebook is based on one in the [spreadsheet modeling textbooks I've used in my classes since 2001](https://host.kelley.iu.edu/albrightbooks/). In both books, they introduce the "Walton Bookstore" problem in the chapter on Monte-Carlo simulation. Here is the basic problem (with a few modifications):

\begin{itemize}
	\item we have to place an order for a perishable product (e.g. a calendar),
	\item there is a known unit cost for each one ordered,
	\item we have a known selling price,
	\item demand is uncertain but we can model it with some simple probability distribution,
	\item for each unsold item, we can get a partial refund of our unit cost,
	\item we need to select the order quantity for our one order for the year; orders can only be in multiples of 25.
\end{itemize}

@@

Then students are given some skeleton code and must finish the expressions for computing key intermediate outputs (\code{order_cost, sales_revenue, refund_revenue}) and the final output, \code{profit}. Answers are provided at the bottom of the notebook so that students can attempt to finish the code but have a resource to correctly complete the code and move on. The screencast associated with the notebook also shows me completing the code. Using skeleton code in this way has worked well both for face to face lab based versions of this type of course as well as for online asynchronous versions.

@@

At this point we discuss a fundamental difference between the computing model of Excel and that of Jupyter notebooks -- while spreadsheets respond automatically to changes, Jupyter notebooks require rerunning all code cells that include or follow the changed code.

@@

Since order quantity is the key decision variable, we might want to see how profit changes for different order quantities in this first simplified model. In Excel, the Data Table tool provides an easy way to do this and is a staple of spreadsheet based management science textbooks. In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed. In Python, we can rely on the vectorized behavior of the language. A range of order quantities is created as a NumPy array that can be used directly to compute vectors of all of the intermediate and final output variables. Much like Data Tables, vectorized computations avoid having to explicitly iterate, or loop, through a collection of input values.

A natural next step is to do the equivalent of an Excel 2-way Data Table. By creating a profit function that takes all of the base inputs as arguments, we can construct a list comprehension that does what an Excel 2-Way Data Table does. Even better, this actually allows us to do the equivalent of an n-Way Data Table with an arbitrary number of outputs - something Excel does not have. This example provides an effective way to show the power of creating reusable functions and Python's handy list comprehension tool.

@@

It is not clear whether an OO or a non-OO approach will end up making the most sense for doing spreadsheet type modeling. This uncertainty is made explicit in the notebook and students are led on a bit of a journey of discovery as we explore different software designs for this modeling problem. I purposefully designed the notebooks in this way to show students that the path to a solution is not necessarily linear and that giving yourself the freedom to explore alternatives, many of which will be dead ends, can lead to a rich learning experience. It is also more realistic in the sense that this is how most real problems are solved. It is analogous to the idea that mathematical proofs usually do not reflect the myriad of failed approaches that were tried but that ultimately contributed to the final product.

@@

For our initial design of a \code{BookstoreModel} class we make all of the base inputs class attributes (properties) and then add method attributes to compute outputs such as costs, revenues and profits. Students can easily see the mapping between things in the initial procedural model and this new object oriented version of the model. I highlight several important concepts and syntactical details of doing OO programming Python. Again, the goal is to weave in these more advanced Python programming concepts within the context of a very familiar modeling problem. This section of the notebook ends with us using our \code{BookstoreModel} class to create a model object with all of its base inputs instantiated with values. Before moving on students are presented with a few challenges involving enhancements to the \code{BookstoreModel} class. Interspersing short coding challenges throughout the notebook gives students a chance to test themselves on their understanding and to get some coding practice.

Then a few design dilemmas are posed and some non-working approaches illustrated as we try to decide on how to implement an n-way data table function for this new object oriented model. These meta reflections are intentionally included to better mimic actual software development and reinforce the idea that software development is much more than simply writing code to implement a perfectly thought out design. We forge ahead with some preliminary ideas and are confronted with an intermediate problem related to generating a list of dictionaries that represent combinations of inputs (i.e. scenarios) to evaluate by our n-way data table function. This provides a great opportunity to remind students of one of the great strengths of the open source software ecosystem - we can leverage work done in other packages and even look at actual source code to see how something was implemented. In this case, the well known scikit-learn package has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely. This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others -- most data science related Python packages are object oriented in nature.

This first notebook draws to a close with the creation of an \code{data_table} function that allows us to do sensitivity analysis with the objected oriented model. Unlike Excel's Data Table tool, this \code{data_table} function can handle an arbitrary number of inputs and outputs. The payoff is the creation of the following faceted plot showing how profit varies for different order quantities and demand levels. We recap the main things learned in this notebook and prepare to add goal seeking capability to our model.

@@

No self respecting modeling tool would be complete without goal seeking capability to do things such as finding the break even point in the bookstore model. For many students, Excel's Goal Seek tool can seem almost magical when encountered for the first time. Even after they get a sense of what it is doing, understanding the importance of the initial guess and how it can lead to Goal Seek reporting different solutions, is not very transparent. With Python, students can gain a much better understanding of how tools like Goal Seek really work.

Before trying to build our \code{goal_seek} function we take a brief detour into basic root finding algorithms through packages like SciPy as well as in standalone Python scripts that are widely available in the Python data science ecosystem (refs). We compare the output of Excel's Goal Seek with different starting values to the values obtained by different root finding algorithms implemented in Python. The importance of starting values and also the range of different root finding algorithms available is often surprising to business students who have spent most of their life in Excel.

A \code{} function is created that implements a simple bisection search and we use it to find the break even demand point in our bookstore model. With \code{data_table} and \code{goal_seek} functions implemented, we are ready to move on to doing Monte-Carlo simulation.

@@

\begin{quote}
	content...

@@

How aap course fits in relation to pcda course. Could also be second part of a fully Python based course. Builds on basic intro Python concepts and skills.

The Quant Econ website is a good example of another use of Python for typical Excel things.

Everything discussed in this article is available from the publicly accessible course website.

>>>>

% User added usepackage commands from Jupyter notebook preamble
\usepackage{tcolorbox}
%\usepackage{graphicx}
%\usepackage{float}
%\usepackage{xcolor} % Allow colors to be defined
\usepackage{enumerate} % Needed for markdown enumerations to work
\usepackage{geometry} % Used to adjust the document margins
%\usepackage{amsmath} % Equations
%\usepackage{amssymb} % Equations
\usepackage{textcomp} % defines textquotesingle
% Hack from http://tex.stackexchange.com/a/47451/13684:
\AtBeginDocument{%
	\def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
}
\usepackage{upquote} % Upright quotes for verbatim code
\usepackage{eurosym} % defines \euro

\usepackage{fancyvrb} % verbatim replacement that allows latex

% Private macros for Jupyter notebook styling

% Colors for the hyperref package
\definecolor{urlcolor}{rgb}{0,.145,.698}
\definecolor{linkcolor}{rgb}{.71,0.21,0.01}
\definecolor{citecolor}{rgb}{.12,.54,.11}

% ANSI colors
\definecolor{ansi-black}{HTML}{3E424D}
\definecolor{ansi-black-intense}{HTML}{282C36}
\definecolor{ansi-red}{HTML}{E75C58}
\definecolor{ansi-red-intense}{HTML}{B22B31}
\definecolor{ansi-green}{HTML}{00A250}
\definecolor{ansi-green-intense}{HTML}{007427}
\definecolor{ansi-yellow}{HTML}{DDB62B}
\definecolor{ansi-yellow-intense}{HTML}{B27D12}
\definecolor{ansi-blue}{HTML}{208FFB}
\definecolor{ansi-blue-intense}{HTML}{0065CA}
\definecolor{ansi-magenta}{HTML}{D160C4}
\definecolor{ansi-magenta-intense}{HTML}{A03196}
\definecolor{ansi-cyan}{HTML}{60C6C8}
\definecolor{ansi-cyan-intense}{HTML}{258F8F}
\definecolor{ansi-white}{HTML}{C5C1B4}
\definecolor{ansi-white-intense}{HTML}{A1A6B2}
\definecolor{ansi-default-inverse-fg}{HTML}{FFFFFF}
\definecolor{ansi-default-inverse-bg}{HTML}{000000}

% Pygments definitions
\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
	\let\PY@ul=\relax \let\PY@tc=\relax%
	\let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
	\PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
				\PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\@namedef{PY@tok@w}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\@namedef{PY@tok@c}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@cp}{\def\PY@tc##1{\textcolor[rgb]{0.61,0.40,0.00}{##1}}}
\@namedef{PY@tok@k}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kp}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kt}{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\@namedef{PY@tok@o}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@ow}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\@namedef{PY@tok@nb}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@nf}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\@namedef{PY@tok@nc}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\@namedef{PY@tok@nn}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\@namedef{PY@tok@ne}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.80,0.25,0.22}{##1}}}
\@namedef{PY@tok@nv}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@no}{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\@namedef{PY@tok@nl}{\def\PY@tc##1{\textcolor[rgb]{0.46,0.46,0.00}{##1}}}
\@namedef{PY@tok@ni}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.44,0.44,0.44}{##1}}}
\@namedef{PY@tok@na}{\def\PY@tc##1{\textcolor[rgb]{0.41,0.47,0.13}{##1}}}
\@namedef{PY@tok@nt}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@nd}{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\@namedef{PY@tok@s}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@sd}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@si}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.64,0.35,0.47}{##1}}}
\@namedef{PY@tok@se}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.36,0.12}{##1}}}
\@namedef{PY@tok@sr}{\def\PY@tc##1{\textcolor[rgb]{0.64,0.35,0.47}{##1}}}
\@namedef{PY@tok@ss}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@sx}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@m}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@gh}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\@namedef{PY@tok@gu}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\@namedef{PY@tok@gd}{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\@namedef{PY@tok@gi}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.52,0.00}{##1}}}
\@namedef{PY@tok@gr}{\def\PY@tc##1{\textcolor[rgb]{0.89,0.00,0.00}{##1}}}
\@namedef{PY@tok@ge}{\let\PY@it=\textit}
\@namedef{PY@tok@gs}{\let\PY@bf=\textbf}
\@namedef{PY@tok@gp}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\@namedef{PY@tok@go}{\def\PY@tc##1{\textcolor[rgb]{0.44,0.44,0.44}{##1}}}
\@namedef{PY@tok@gt}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\@namedef{PY@tok@err}{\def\PY@bc##1{{\setlength{\fboxsep}{\string -\fboxrule}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}}
\@namedef{PY@tok@kc}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kd}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kn}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kr}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@bp}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@fm}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\@namedef{PY@tok@vc}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@vg}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@vi}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@vm}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@sa}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@sb}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@sc}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@dl}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@s2}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@sh}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@s1}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@mb}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@mf}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@mh}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@mi}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@il}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@mo}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@ch}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@cm}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@cpf}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@c1}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@cs}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % Define a nice break command that doesn't care if a line doesn't already
% exist.
\def\br{\hspace*{\fill} \\* }

% For linebreaks inside Verbatim environment from package fancyvrb.
\makeatletter
\newbox\Wrappedcontinuationbox
\newbox\Wrappedvisiblespacebox
\newcommand*\Wrappedvisiblespace {\textcolor{red}{\textvisiblespace}}
\newcommand*\Wrappedcontinuationsymbol {\textcolor{red}{\llap{\tiny$\m@th\hookrightarrow$}}}
\newcommand*\Wrappedcontinuationindent {3ex }
\newcommand*\Wrappedafterbreak {\kern\Wrappedcontinuationindent\copy\Wrappedcontinuationbox}
% Take advantage of the already applied Pygments mark-up to insert
% potential linebreaks for TeX processing.
%        {, <, #, %, $, ' and ": go to next line.
%        _, }, ^, &, >, - and ~: stay at end of broken line.
% Use of \textquotesingle for straight quote.
\newcommand*\Wrappedbreaksatspecials {%
	\def\PYGZus{\discretionary{\char`\_}{\Wrappedafterbreak}{\char`\_}}%
	\def\PYGZob{\discretionary{}{\Wrappedafterbreak\char`\{}{\char`\{}}%
	\def\PYGZcb{\discretionary{\char`\}}{\Wrappedafterbreak}{\char`\}}}%
	\def\PYGZca{\discretionary{\char`\^}{\Wrappedafterbreak}{\char`\^}}%
	\def\PYGZam{\discretionary{\char`\&}{\Wrappedafterbreak}{\char`\&}}%
	\def\PYGZlt{\discretionary{}{\Wrappedafterbreak\char`\<}{\char`\<}}%
	\def\PYGZgt{\discretionary{\char`\>}{\Wrappedafterbreak}{\char`\>}}%
	\def\PYGZsh{\discretionary{}{\Wrappedafterbreak\char`\#}{\char`\#}}%
	\def\PYGZpc{\discretionary{}{\Wrappedafterbreak\char`\%}{\char`\%}}%
	\def\PYGZdl{\discretionary{}{\Wrappedafterbreak\char`\$}{\char`\$}}%
	\def\PYGZhy{\discretionary{\char`\-}{\Wrappedafterbreak}{\char`\-}}%
	\def\PYGZsq{\discretionary{}{\Wrappedafterbreak\textquotesingle}{\textquotesingle}}%
	\def\PYGZdq{\discretionary{}{\Wrappedafterbreak\char`\"}{\char`\"}}%
	\def\PYGZti{\discretionary{\char`\~}{\Wrappedafterbreak}{\char`\~}}%
}
% Some characters . , ; ? ! / are not pygmentized.
% This macro makes them "active" and they will insert potential linebreaks
\newcommand*\Wrappedbreaksatpunct {%
	\lccode`\~`\.\lowercase{\def~}{\discretionary{\hbox{\char`\.}}{\Wrappedafterbreak}{\hbox{\char`\.}}}%
	\lccode`\~`\,\lowercase{\def~}{\discretionary{\hbox{\char`\,}}{\Wrappedafterbreak}{\hbox{\char`\,}}}%
	\lccode`\~`\;\lowercase{\def~}{\discretionary{\hbox{\char`\;}}{\Wrappedafterbreak}{\hbox{\char`\;}}}%
	\lccode`\~`\:\lowercase{\def~}{\discretionary{\hbox{\char`\:}}{\Wrappedafterbreak}{\hbox{\char`\:}}}%
	\lccode`\~`\?\lowercase{\def~}{\discretionary{\hbox{\char`\?}}{\Wrappedafterbreak}{\hbox{\char`\?}}}%
	\lccode`\~`\!\lowercase{\def~}{\discretionary{\hbox{\char`\!}}{\Wrappedafterbreak}{\hbox{\char`\!}}}%
	\lccode`\~`\/\lowercase{\def~}{\discretionary{\hbox{\char`\/}}{\Wrappedafterbreak}{\hbox{\char`\/}}}%
	\catcode`\.\active
	\catcode`\,\active
	\catcode`\;\active
	\catcode`\:\active
	\catcode`\?\active
	\catcode`\!\active
	\catcode`\/\active
	\lccode`\~`\~
}
\makeatother

\let\OriginalVerbatim=\Verbatim
\makeatletter
\renewcommand{\Verbatim}[1][1]{%
	%\parskip\z@skip
	\sbox\Wrappedcontinuationbox {\Wrappedcontinuationsymbol}%
	\sbox\Wrappedvisiblespacebox {\FV@SetupFont\Wrappedvisiblespace}%
	\def\FancyVerbFormatLine ##1{\hsize\linewidth
		\vtop{\raggedright\hyphenpenalty\z@\exhyphenpenalty\z@
			\doublehyphendemerits\z@\finalhyphendemerits\z@
			\strut ##1\strut}%
	}%
	% If the linebreak is at a space, the latter will be displayed as visible
	% space at end of first line, and a continuation symbol starts next line.
	% Stretch/shrink are however usually zero for typewriter font.
	\def\FV@Space {%
		\nobreak\hskip\z@ plus\fontdimen3\font minus\fontdimen4\font
		\discretionary{\copy\Wrappedvisiblespacebox}{\Wrappedafterbreak}
		{\kern\fontdimen2\font}%
	}%

	% Allow breaks at special characters using \PYG... macros.
	\Wrappedbreaksatspecials
	% Breaks at punctuation characters . , ; ? ! and / need catcode=\active
	\OriginalVerbatim[#1,codes*=\Wrappedbreaksatpunct]%
}
\makeatother

% Exact colors from NB
\definecolor{incolor}{HTML}{303F9F}
\definecolor{outcolor}{HTML}{D84315}
\definecolor{cellborder}{HTML}{CFCFCF}
\definecolor{cellbackground}{HTML}{F7F7F7}

% prompt
\makeatletter
\newcommand{\boxspacing}{\kern\kvtcb@left@rule\kern\kvtcb@boxsep}
\makeatother
\newcommand{\prompt}[4]{
	{\ttfamily\llap{{\color{#2}[#3]:\hspace{3pt}#4}}\vspace{-\baselineskip}}
}

% Prevent overflowing lines due to hard-to-break entities
\sloppy
% Setup hyperref package
\hypersetup{
	breaklinks=true,  % so long urls are correctly broken across lines
	colorlinks=true,
	urlcolor=urlcolor,
	linkcolor=linkcolor,
	citecolor=citecolor,
}
% Slightly bigger margins than the latex defaults

\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

@@

Created PCDA course in response. Challenges with finding good textbooks. Developed much of my own curricular materials. Taught using combo of R Markdown documents and Jupyter notebooks. Literate computing \cite{Jupyter_proposal}. Educational narratives that support iterative exploration and development. The ability to weave standard language with formal programming languages is critical to creating the desired educational artifact. Jupyter used by many universities as a key component of their educational technology stack.

PCDA course focused mostly on data wrangling, EDA, statistical and machine learning based predictive models though I did include a Python based simulation of the Monte Hall 3-door problem. In teaching EDA using tools like ggplot2, I would challenge students to think about how they would go about trying to do some of the plots we were making in Excel. There is also some limited coverage of using Python to open Excel files and do some sheet manipulation. However, there was no real attempt to do the kind of modeling that was the focus of textbooks like PMS or DADM.

@@

Jupyter notebooks for teaching data science are an established delivery mechanism. Computational notebook allowing mixing of explanatory text, written in Markdown with executable code (and its output) in Python or a number of other languages. Literate programming. Can also include equations, visualizations, other multi-media. It's free, open source, runs on all platforms, huge ecosystem, used in industry, used widely in education.

@@

There are two overarching learning objectives for this first set of Jupyter notebooks. Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet - model building and sensitivity analysis. At the same time, students encounter, as they are needed, more advanced Python programming material. In this way, the computational problem to be solved drives the introduction of more advanced Python concepts and techniques instead of such things being presented in a vacuum. I make it a point to try to convey the notion that there are often multiple possible approaches by embarking on one path and there can be value to starting over and doing things in a different way. Each notebook is accompanied by videos that walk the students through the material and act as a guided tour with challenges to overcome along the way. The screenshot in \ref{fig:activities} shows the portion of the course web page that provides the high level roadmap through the notebooks and associated screencasts.

@@

The stage is set with an overview of a typical spreadsheet based model that most of the students have seen before.

@@

With this as motivation, we then introduce a similar modeling problem. The example used throughout the rest of the notebook is based on one in the [spreadsheet modeling textbooks I've used in my classes since 2001](https://host.kelley.iu.edu/albrightbooks/). In both books, they introduce the "Walton Bookstore" problem in the chapter on Monte-Carlo simulation. Here is the basic problem (with a few modifications):

\begin{itemize}
	\item we have to place an order for a perishable product (e.g. a calendar),
	\item there is a known unit cost for each unit ordered,
	\item we have a known and fixed selling price,
	\item demand is uncertain but we can model it with some simple probability distribution,
	\item for each unsold item, we can get a partial refund of our unit cost,
	\item we need to select the order quantity for our one order for the year; orders can only be in multiples of 25.
\end{itemize}

@@

\begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
	%\prompt{In}{incolor}{3}{\boxspacing}
	\begin{Verbatim}[commandchars=\\\{\}]
	\PY{c+c1}{\PYZsh{} Base inputs}
	\PY{n}{unit\PYZus{}cost} \PY{o}{=} \PY{l+m+mf}{7.50}
	\PY{n}{selling\PYZus{}price} \PY{o}{=} \PY{l+m+mf}{10.00}
	\PY{n}{unit\PYZus{}refund} \PY{o}{=} \PY{l+m+mf}{2.50}
	\end{Verbatim}
\end{tcolorbox}

Students are given skeleton code as scaffolding and must finish the expressions for computing key intermediate outputs (\code{order\_cost, sales\_revenue, refund\_revenue}) and the final output, \code{profit}. Answers are provided at the bottom of the notebook so that students can attempt to finish the code but have a resource to correctly complete the code and move on. The screencast associated with the notebook also shows me completing the code in a subsequent code cell. Using skeleton code in this way has worked well both for lab based versions of this type of course as well as for online asynchronous versions.

@@

At this point we discuss a fundamental difference between the computing model of Excel and that of Jupyter notebooks -- while spreadsheets respond automatically to changes, Jupyter notebooks require rerunning all code cells that include or follow the changed code. This difference in behavior between Jupyter notebooks and spreadsheets might be one of the most important concepts for students to internalize. A well known challenge associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors \cite{bibid}. This point along with strategies to avoid it are discussed during the first week of the course which focuses on notebook computing best practices as well as an historical look at the use of computational notebooks \cite{bibid}.

@@

Since order quantity is the key decision variable, we might want to see how profit changes for different order quantities. In Excel, the Data Table tool provides an easy way to do this and is a staple of spreadsheet based management science textbooks. In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed. In Python, we can rely on the vectorized behavior of the language \cite{bibid}. A range of order quantities is created as a NumPy array that can be used directly to compute vectors of all of the intermediate and final output variables. Much like Data Tables, vectorized computations avoid having to explicitly iterate, or loop, through a collection of input values.

A natural next step is to do the equivalent of an Excel 2-way Data Table. By creating a profit function that takes all of the base inputs as arguments, we can construct a list comprehension that does what an Excel 2-Way Data Table does. Even better, this actually allows us to do the equivalent of an n-Way Data Table with an arbitrary number of outputs - something Excel does not have. This example provides an effective way to show the power of creating reusable functions and Python's handy list comprehension construct.

@@

It is not clear whether an object oriented (OO) or a non-OO approach will end up making the most sense for doing spreadsheet type modeling. This uncertainty is made explicit in the notebook and students are led on a bit of a journey of discovery as we explore different software designs for this modeling problem. I purposefully designed the notebooks in this way to show students that the path to a solution is not necessarily linear and that giving yourself the freedom to explore alternatives, many of which will be dead ends, can lead to a rich learning experience. It is also more realistic in the sense that this is how most real problems are solved. It is analogous to the idea that mathematical proofs usually do not reflect the myriad of failed approaches that were tried but that ultimately contributed to the final product.

@@

For our initial design of a \code{BookstoreModel} class we make all of the base inputs class attributes (properties) and then add method attributes to compute outputs such as costs, revenues and profits. Students can easily see the mapping between things in the initial procedural model and this new object oriented version of the model. I highlight several important concepts and syntactical details of doing object oriented programming Python. Again, the goal is to weave in these more advanced Python programming concepts within the context of a very familiar modeling problem. This section of the notebook ends with us using our \code{BookstoreModel} class to create a model object with all of its base inputs instantiated with values. Before moving on students are presented with a few challenges involving enhancements to the \code{BookstoreModel} class. Interspersing short coding challenges throughout the notebook gives students a chance to test themselves on their understanding and to get some coding practice.

Then a few design dilemmas are posed and some non-working approaches illustrated as we try to decide on how to implement an n-way data table function for this new object oriented model. These meta reflections are intentionally included to better mimic actual software development and reinforce the idea that software development is much more than simply writing code to implement a perfectly thought out design. We forge ahead with some preliminary ideas and are confronted with an intermediate problem related to generating a list of dictionaries that represent combinations of inputs (i.e.\ scenarios) to evaluate by our n-way data table function. This provides a great opportunity to remind students of one of the great strengths of the open source software ecosystem - we can leverage work done in other packages and even look at actual source code to see how something was implemented. In this case, the well known scikit-learn package \cite{bibid} has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely. This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others as most data science related Python packages are object oriented in nature.

This first notebook draws to a close with the creation of a \code{data\_table} function that allows us to do sensitivity analysis with the objected oriented model. Unlike Excel's Data Table tool, this \code{data\_table} function can handle an arbitrary number of inputs and outputs. The payoff is the creation of the following faceted plot showing how profit varies for different order quantities and demand levels. We recap the main things learned in this notebook and prepare to add goal seeking capability to our model.

@@

No self respecting modeling tool would be complete without goal seeking capability to do things such as finding the break even point in the bookstore model. For many students, Excel's Goal Seek tool can seem almost magical when encountered for the first time. Even after they get a sense of what it is doing, understanding the importance of the initial solution guess and how it can lead to Goal Seek reporting different solutions, is not very transparent. With Python, students can gain a much better understanding of how tools like Goal Seek really work.

Before trying to build our \code{goal\_seek} function we take a brief detour into basic root finding algorithms through packages like SciPy \cite{bibid} as well as in standalone Python scripts that are widely available in the Python data science ecosystem \cite{bibid}. We compare the output of Excel's Goal Seek with different starting values to the values obtained by different root finding algorithms implemented in Python. The importance of starting values and also the range of different root finding algorithms available is often surprising to business students who have spent most of their analytical life in Excel.

A \code{goal\_seek} function is created that implements a simple bisection search and we use it to find the break even demand point in our bookstore model. With \code{data\_table} and \code{goal\_seek} functions implemented, we are ready to move on to doing Monte-Carlo simulation.

@@

\begin{quote}

@@
